\documentclass{sig-alternate}

\usepackage{hyperref}
\usepackage{macros}
\usepackage{amscd}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}

\numberofauthors{3}
\author{
  \alignauthor Luca De Feo\\
  \affaddr{Universit\'e Paris Saclay -- UVSQ, LMV}\\
  \email{luca.de-feo@uvsq.fr}
  \alignauthor Hugues Randriam
  \affaddr{LTCI, T\'el\'ecom ParisTech}\\
  \email{randriam@enst.fr}
  \alignauthor \'Edouard Rousseau
  \affaddr{LTCI, T\'el\'ecom ParisTech, UVSQ, LMV}\\
  \email{erousseau@enst.fr}
}

\title{Efficient lattices of compatibly embedded finite fields}

\begin{document}

\maketitle

\begin{abstract}
  Lattices of compatibly embedded finite fields are useful in computer
  algebra systems for managing many extensions of a finite field
  $\F_p$ at once. %
  They can also be used to represent the algebraic closure $\bar\F_p$,
  and to represent all finite fields in a standard manner.

  The most well known constructions are Conway polynomials, and the
  Bosma-Cannon-Steel framework used in Magma. %
  They all have drawbacks: Conway polynomials are extremely expensive
  to compute, while the Bosma-Cannon-Steel framework does not provide
  a standard way to represent finite fields, and becomes inefficient
  as the number of extensions grows.

  In this work, leveraging the theory of the Lenstra-Allombert
  isomorphism algorithm, we generalize at the same time Conway
  polynomials and the Bosma-Cannon-Steel framework, obtaining the
  first construction with quadratic complexity in the extension
  degree, and linear complexity in the number of extensions. %
  We also provide an implementation in C/Flint/Julia/Nemo, showing
  that our construction is practical.
\end{abstract}

\category{F.2.1}{Theory of computation}{Analysis of algorithms and problem complexity}[Computations in finite fields]
\category{G.4}{Mathematics of computing}{Mathematical software}
\terms{Algorithms,Theory}
\keywords{Finite fields, field extensions, Conway polynomials.}

\section{Introduction}
\label{sec:introduction}

Computer algebra systems (CAS) are often faced with the problem of
constructing several extensions of a finite field $\F_p$ in a
\emph{compatible} way, i.e., such that the (subfield) inclusion
lattice of the given extensions can be computed and evaluated
efficiently.

Concretely, what is sought is a data structure $\Lambda$ to represent
arbitrary collections of extensions of $\F_{p}$, in such a way that
elements of $\F_{p^n}$ are represented in optimal space (i.e., $O(n)$
coefficients), and that arithmetic operations are performed
efficiently (i.e., $O\left(\lcm(n,m)^d\right)$ arithmetic operations
to combine an element of $\F_{p^m}$ and an element of $\F_{p^n}$,
where $d\le 3$ and, possibly, $d=1+\varepsilon$). %
To this end, it is useful to set several sub-goals:

\begin{description}
\item[\emph{Effective embeddings:}] For any pair of extensions
  $k\subset K$ in $\Lambda$, there exists an efficiently computable
  embedding $\phi:k\to K$, and algorithms to evaluate $\phi$ on $k$,
  and the section $\phi^{-1}$ on $K$.
\item[\emph{Compatibility:}] The embeddings are \emph{compatible},
  i.e., for any triple $k\subset K\subset L$ in $\Lambda$, and
  embeddings $\phi:k\to K$, $\psi:K\to L$, $\chi:k\to L$, one has
  $\chi=\psi\circ\phi$.
\item[\emph{Incrementality:}] The data associated with an extension
  (e.g., its irreducible polynomial, change-of-basis matrices, \dots)
  must be computable efficiently and \emph{incrementally}, i.e.,
  adding a new field extension to $\Lambda$ does not require
  recomputing data for all extensions already in $\Lambda$. %
\item[\emph{Uniqueness:}] Any extension of $\F_{p}$ is determined by an
  irreducible polynomial whose definition only depends on the
  characteristic $p$ and the degree of the extension. %
\item[\emph{Generality:}] Extensions of $\F_{p}$ can be represented by
  arbitrary irreducible polynomials.
\end{description}

Some goals, such as incrementality, uniqueness and generality are
optional, and it is obvious that uniqueness and generality are even in
conflict with each other. %
An incremental data structure can be used to effectively represent an
algebraic closure $\bar\F_{p}$, with new finite extensions built on the
fly as they are needed. %
Uniqueness is useful for defining field elements in a standard way,
portable between different CAS, while generality is useful in a
context where the user is left with the freedom of choosing the
defining polynomials. %
Note that any solution can be made unique by replacing all random
choices with pseudo-random ones, however one is usually interested in
uniqueness solutions that have a simple mathematical description. %
Also, any solution can be made general by means of an isomorphism
algorithm~\cite{LenstraJr91,Allombert02,rains2008,brieulle2018computing,narayanan2016fast}. %
Other optional goals, such as computing normal bases or evaluating
Frobenius morphisms, may be added to the list, however they are out of
the scope of this work.

\paragraph{Previous work}
The first and most well known solution is the family of Conway
polynomials~\cite{Nickel1988,heath+loehr99}, first adopted in
GAP~\cite{GAP4}, and then also by Magma~\cite{MAGMA} and
Sage~\cite{Sage}. %
Conway polynomials yield uniqueness, however computing them requires
exponential time, thus incrementality is only available at a
prohibitive cost; for this reason, they are usually pre-computed and
tabulated up to some bound.

The first to show the existence of a (incremental, general) data
structure computable in deterministic polynomial time was
Lenstra~\cite{LenstraJr91}, who proved that, besides the problem of
finding irreducible polynomials, any other question is amenable to
linear algebra. %
Subsequent work of Lenstra and de
Smit~\cite{lenstra+desmit08-stdmodels} tackled the uniqueness problem,
albeit only from a theoretical point of view. %

In practice, randomized algorithms are good enough for a CAS, then
polynomial factorization and basic linear algebra provide an easy
(incremental, general) solution, that was first analyzed by Bosma,
Cannon and Steel~\cite{bosma+cannon+steel97}, and is currently used in
Magma. %

All solutions presented so far have superquadratic complexity, i.e.,
$d>2$. %
Recent work on embedding algorithms~\cite{DoSc12,DeDoSc13,DeDoSc2014}
yields subquadratic (more precisely, $d\le 1.5$) solutions for
specially constructed (non-unique, non-general) families of
irreducible polynomials, and even quasi-optimal ones (i.e.,
$d=1+\varepsilon$) if a quasi-linear modular composition algorithm is
available. %
However these constructions involve counting points of random elliptic
curves over finite fields, and have thus a rather high polynomial
dependency in $\log p$; for this reason, they are usually considered
practical only for relatively small characteristic.


\paragraph{Our contribution}
In this work we present the first
quadratic, % TODO: let's hope this is true...
incremental, general and/or unique % TODO: also this...
solution for lattices of compatibly embedded finite fields. %
Our starting point is Allombert's~\cite{Allombert02} and
subsequent~\cite{brieulle2018computing} improvements to Lenstra's
isomorphism algorithm. %
Plugging them in the Bosma-Cannon-Steel framework immediately produces
a quadratic, incremental general solution, however we go much
further. %
Indeed, we show that the compatibility requirement can be taken a step
further by constructing a lattice of $\F_{p}$-algebras with a
distinguished element, which is a byproduct of the Lenstra-Allombert
algorithm.

The advantages of our construction over a naive combination of the
Lenstra-Allombert algorithm and the Bosma-Cannon-Steel framework are
multiple. %
Storage drops from quadratic to linear in the number of extensions
stored in $\Lambda$, and the cost of adding a new extension to
$\Lambda$ drops similarly. %
This yields a significant speedup in practice, as we show in our
implementation (see Section~\ref{sec:implementation}).

Our $\F_{p}$-algebras are constructed by tensoring element-wise an
arbitrary lattice of extensions of $\F_{p}$ with a lattice of compatible
roots of unity. %
One of the possible ways to construct the second lattice is by using
Conway polynomials; this way, our construction can be seen as a
generalization of Conway polynomials, permitting to represent
(uniquely) an exponentially larger set of finite fields than with
Conway polynomials alone. %
Since the storage overhead for the second lattice is only logarithmic
in the degree of the extension fields, the high cost for computing
Conway polynomials is absorbed in the total complexity.

The main drawback of our method is that, in some instances, combining
elements of $\F_{p^m}$ and $\F_{p^n}$ requires constructing a field
extension of degree larger than $\lcm(m,n)$. %
While this does not affect the asymptotic complexity of our
construction, % TODO: doesn't it?
it can in practice make some computations visibly slower.


\paragraph{Organisation}
The presentation is structured as follows. %
In Section~\ref{sec:conway} we review some basic facts on roots of
unity and Conway polynomials. %
In Section~\ref{sec:lenstra} we review the Lenstra-Allombert algorithm
and we define and study \emph{Kummer algebras},
the main ingredient to our construction. %
In Section~\ref{sec:compatibleH90} we introduce a notion of compatibility
for solutions of Hilbert~90 in Kummer algebras, that provides standard
defining polynomials for finite fields.
Then in Section~\ref{sec:construction} we again use these compatible solutions
to construct standard compatible embeddings between finite fields,
from which a lattice can be incrementally constucted.
Finally, in Section~\ref{sec:implementation} we present our
implementation and compare it to available alternatives.


\section{Preliminaries}
\label{sec:conway}

\paragraph{Fundamental algorithms and complexities}
Throughout this paper we let $\F_p$ be a prime field, and we measure
time complexities as a number of arithmetic operations $(+,\times,/)$
over $\F_p$, and storage as a number of elements of $\F_p$. %
We let $\M(m)$ denote the number of operations required to multiply
two polynomials with coefficients in $\F_p$ of degree at most $m$, and
adopt the usual super-linearity assumptions on the function $\M$
(see~\cite[Ch.~8.3]{vzGG}).

Any finite extension $\F_{p^m}$ can be represented as the quotient of
$\F_p[X]$ by an irreducible polynomial of degree $m$. %
The algorithms we present in the next sections need not assume any
particular representation for finite fields, however when analyzing
their complexities we will assume this representation. %
Then, multiplications in $\F_{p^m}$ can be carried out using
$O(\M(m))$ operations, and inversions using $O(\M(m)\log(m))$. %
In this work we will also need to perform computations in algebras
$\F_{p^m}\otimes\F_{p^n}$; representing them as quotients of a
bivariate polynomial ring, we can multiply elements using $O(\M(mn))$
operations.

We will extensively use a few standard routines, that we recall
briefly. %
Brent and Kung's algorithm~\cite{brent+kung} computes the
\textbf{modular composition} $f(g)\bmod h$ of three polynomials
$f,g,h\in\F_p[X]$ of degree at most $m$ using
$O\bigl(m^{(\omega+1)/2)}\bigr)$ operations, where $\omega$ is the
\emph{exponent of linear algebra} over $\F_p$. %
The Kedlaya--Umans algorithm~\cite{KeUm11} solves the same problem,
and has better complexity in the binary RAM model, however it is
widely considered impractical, we shall thus not consider it in our
complexity estimates.

By applying \emph{transposition
  techniques}~\cite{burgisser+clausen-shokrollahi,bostan+lecerf+schost:tellegen}
to Brent and Kung's algorithm, Shoup~\cite{shoup94,shoup99} derived an
algorithm to compute \textbf{minimal polynomials} of arbitrary
elements of $\F_{p^m}$, having the same complexity
$O\bigl(m^{(\omega+1)/2)}\bigr)$. %
Kedlaya and Umans' improvements also apply to Shoup's minimal
polynomial algorithm, with the same practical limitations.

Shoup's techniques can also be applied to \textbf{evaluate embeddings
  of finite fields}. %
Let $\F_{p^\ell},\F_{p^m}$ be a pair of finite fields related by an
embedding $\phi:\F_{p^\ell}\hookrightarrow\F_{p^m}$, and let a
generator $\alpha_\ell$ of $\F_{p^\ell}$ and its image
$\phi(\alpha_\ell)$ in $\F_{p^m}$ be given. %
Given these data, for any element $x\in\F_{p^\ell}$ it is possible to
compute its image $\phi(x)$ using $O\bigl(m^{(\omega+1)/2)}\bigr)$
operations; similarly, given an element $y=\phi(x)$ in $\F_{p^m}$, it
is possible to recover $x$ in the same asymptotic number of
operations. %
The relevant algorithms are reviewed in~\cite[Sec.~6]{ffisom-long};
note that, for specially constructed generators $\alpha_\ell$, more
efficient algorithms may
exist~\cite{df+schost12,DoSc12,DeDoSc13,DeDoSc2014}.

The present work focuses on algorithms to \textbf{compute embeddings
  of finite fields}, i.e., algorithms that, given finite fields
$\F_{p^\ell},\F_{p^m}$ with $\ell\,|\,m$, find an embedding
$\phi:\F_{p^\ell}\hookrightarrow\F_{p^m}$, a generator $\alpha_\ell$
of $\F_{p^\ell}$, and its image $\phi(\alpha_\ell)$. %
An extensive review of known algorithms is given
in~\cite{brieulle2018computing}; here we shall only be interested in
the Lenstra--Allombert isomorphism
algorithm~\cite{LenstraJr91,Allombert02}, and its adaptation to
compatible lattices of finite fields.


\paragraph{Conway polynomials and cyclotomic lattices}
The algorithms of the next sections will be dependent on the
availability of a compatible lattice of finite field elements. %
Let $I$ be a subset of $\N\setminus p\N$, we call \emph{cyclotomic
  lattice} a function $\zeta:I\to\bar\F_p$ such that
\begin{align}
  \zeta(1) &= 1,\\
  \label{eq:cyclotomic}
  \zeta(\ell) &= \zeta(m)^{m/l} &\text{for any $\ell\,|\,m$,}\\
  \zeta(\ell) &= \zeta(m)      &\text{iff $\ell=m$.}
\end{align}
Said otherwise, a cyclotomic lattice is a collection of primitive
roots of unity in $\bar\F_p$ with a compatibility condition. %
In the next sections, we shall assume that such a lattice is given in
advance, for a set $I$ large enough, and we shall write $\zeta_\ell$ for
$\zeta(\ell)$.

If $I$ is a finite set of indices, there is an easy randomized
algorithm to construct a cyclotomic lattice: compute
$n=\lcm_{i\in I}(i)$, construct the smallest field $\F_{p^a}$ such
that $n$ divides $p^a-1$, take a random $x^{(p^a-1)/n}\in\F_{p^a}$ and
test that it has multiplicative order $n$; all other roots in the
lattice are constructed as powers of this element.

Nevertheless, the most useful cyclotomic lattices are those where $I$
is the whole set $\N\setminus p\N$. % ai-je eu raison de corriger ?
This bears the question of how to represent $\bar\F_p$. %
It may seem odd to ask for a representation of $\bar\F_p$ as a
prerequisite for a construction whose goal is precisely to represent
$\bar\F_p$. %
However we shall see that a relatively small cyclotomic sub-lattice is
enough to construct a much larger lattice of compatibly embedded
finite fields; it thus makes sense to assume that a cyclotomic lattice
is available, if it can be computed \emph{incrementally}.

\emph{Conway polynomials}~\cite{Nickel1988} offer a classic example of
cyclotomic lattice. %
The $m$-th Conway polynomial $C_m\in\F_p[X]$ is defined as the
\emph{lexicographically smallest} monic irreducible polynomial of
degree $m$ that is also \emph{primitive} (i.e., its roots generate
$\F_{p^n}^\times$) and \emph{norm compatible} (i.e,
$$C_\ell\Bigl(X^{\frac{p^m-1}{p^\ell-1}}\Bigr) = 0 \mod C_m$$
whenever $\ell$ divides $m$). %
The cyclotomic lattice is defined by letting $\zeta(p^m-1)$ be the
image of $X$ in $\F_{p^m}=\F_p[X]/C_m$ for any $m$, and
using~\eqref{eq:cyclotomic} to define $\zeta$ on all other indices. %
Note that, using Conway polynomials, each value of $\zeta$ is
represented in the smallest field containing it.

The best algorithm to compute Conway polynomials has exponential
complexity~\cite{heath+loehr99}, thus they are usually precomputed and
tabulated up to a certain bound. %
Most computer algebra systems switch to other ways of representing
finite fields when the tables of Conway polynomials are not enough. %
A notable exception is SageMath~\cite{Sage} (since version
5.13~\cite{Roe2013}), that relaxes the ``lexicographically first''
requirement and defines \emph{pseudo-Conway polynomials}; although
easier to compute in practice, their computation still requires an
exponential amount of work.

Other ways to construct cyclotomic lattices are possible. %
One may, for example, factor cyclotomic polynomials over $\F_p$, being
careful to maintain compatibility. %
In the next sections we shall not suppose any cyclotomic lattice
construction in particular, and simply assume that we are given a
collection of compatible roots of unity.
That is, formally, we are given a system $\system=\{(K_\ell,\zeta_\ell)\}_{\ell\geq1}$
where $K_\ell$ is an explicitely represented finite extension of $\F_p$,
and $\zeta_\ell\in K_\ell$ a generating element that is also a primitive $\ell$-th root
of unity,
so \[ K_\ell=\F_p(\zeta_\ell),\quad(\zeta_\ell)^\ell=1, \]
together with
explicit embeddings
\[
\begin{array}{cccc}
  \embedcyc{\ell}{m}: & K_\ell & \hookrightarrow & K_m \\
  & \zeta_\ell & \mapsto & (\zeta_{m})^{\frac{m}{\ell}}
\end{array}
\]
whenever $l\,|\,m$.

\section{The Lenstra-Allombert embedding algorithm}
\label{sec:lenstra}

We now review the theory behind Allombert's
adaptation~\cite{Allombert02} of Lenstra's isomorphism
algorithm\cite{LenstraJr91}. %
This will be our stepping stone towards the definition of some
\emph{standard} elements in field extensions of $\F_p$, with an
effective compatibility condition.

The main ingredient of the algorithm is an extension
of Kummer theory.
On any finite extension of $\F_{p}$, denote by $ \sigma:x\mapsto x^p $
the Frobenius automorphism.
Let $\ell$ be an integer not divisible by $p$.
Then $\sigma$ acting on $\F_{p^\ell}$ can be seen as a $\F_{p}$-linear endomorphism,
with minimal polynomial $T^\ell-1$, which is separable but not necessarily split,
i.e. $\F_{p^\ell}$ is not necessarily a Kummer extension of $\F_{p}$.
We make an extension of scalars and work in the
\emph{Kummer algebra of degree~$\ell$}:
\[
  A_\ell = \F_{p^\ell}\otimes\F_{p}(\zeta_\ell),
\]
where $\otimes$ is tensor product over $\F_{p}$, and $\zeta_\ell$ is a primitive $\ell$-th root of unity, taken from a system of such compatible roots. We
call $\F_{p}(\zeta_\ell)$ the \emph{field of scalars} of $A_\ell$, and
we define the \emph{level} of $A_\ell$ as
\[
  \nu(\ell) = \omega_{(\mathbb{Z}/\ell\mathbb{Z})^\times}(p) = [\F_{p}(\zeta_\ell):\F_{p}],
\]
that is, the degree of its field of scalars.

Now $\sigma\otimes1$ is a $1\otimes\F_{p}(\zeta_\ell)$-linear endomorphism of $A_\ell$
with $\ell$ distinct eigenvalues, namely the powers of $1\otimes\zeta_\ell$.
Thus, if $\eta=\zeta_\ell^i$ is any $\ell$-th root of unity in $\F_{p}(\zeta_\ell)$,
the corresponding eigenspace is defined by the \emph{Hilbert 90 equation for $\eta$}:
\begin{equation}
  \tag{H90}
 (\sigma\otimes1)(x) = (1\otimes\eta)x,
  \label{h90}
\end{equation}
which plays the role of $\sigma(x)=\eta x$ in classical Kummer theory.
The solutions of~\eqref{h90} in $A_\ell$ form a
$1\otimes\F_{p}(\zeta_{\ell})$-vector space of dimension $1$,
and if $x$ is such a solution for $\eta$, then $x^j$ is a solution for $\eta^j$.

In particular, let $\alpha_\ell$ be a nonzero solution of~\eqref{h90} for $\zeta_\ell$.
Then $1,\alpha_\ell,\cdots,(\alpha_\ell)^{\ell-1}$ are eigenvectors for distinct eigenvalues and thus
form a basis of $A_\ell$ over $1\otimes\F_{p}(\zeta_{\ell})$.
Likewise
\[ (\alpha_\ell)^\ell = 1\otimes\a_\ell \]
for a certain scalar $\a_\ell\in\F_{p}(\zeta_\ell)$,
the \emph{Kummer constant} of $\alpha_\ell$.
This proves:
\begin{proposition}
\label{alphagen}
Any such $\alpha_\ell$ is a generating element for $A_\ell$ as an algebra over $1\otimes\F_{p}(\zeta_{\ell})$,
inducing an isomorphism
\[ A_\ell\simeq \F_{p}(\zeta_{\ell})[T]/(T^\ell-\a_\ell). \]
\end{proposition}

Since $A_\ell$ is known to be an \'etale algebra, we note
that $\alpha_\ell$ nonzero implies $\a_\ell$ is nonzero, which in turn implies
that $\alpha_\ell$ in \emph{invertible} in $A_\ell$: $\;(\alpha_\ell)^{-1} = (1\otimes\a_\ell^{-1})(\alpha_\ell)^{\ell-1}$.

%We call $\a_\ell$ the \emph{Kummer constant} associated with $\alpha_\ell$.
%In the sequel, we might have to consider various
%solutions of~\eqref{h90}, say $\alpha_\ell,\alpha'_\ell,\alpha_m$.
%With this notation, the index will always implicitely mean that
%these are solutions for the corresponding distinguished primitive root
%of unity, so $\alpha_\ell,\alpha'_\ell$ are solutions for $\zeta_\ell$
%while $\alpha_m$ is solution for $\zeta_m$.
%Then we will also write $\a_\ell,\a'_\ell,\a_m$, which will always
%implicitely denote the associated Kummer constants,
%so e.g. $(\alpha'_\ell)^\ell=1\otimes\a'_\ell$ 
%and $(\alpha_m)^m=1\otimes\a_m$.

We will make frequent use of the following:
\begin{lemma}
\label{FrobFrob}
Let $K,L$ are two finite extensions of $\F_{p}$.
Then for any $\beta\in K\otimes L$,
we have $(\sigma\otimes\sigma)(\beta)=\beta^p$.
\end{lemma}
In this generality we also introduce the following notation: if $\eta\in L$ has degree $d$ over $\F_{p}$,
then any $\beta\in K\otimes\F_{p}(\eta)\subset K\otimes L$ decomposes uniquely as $\beta = \sum_{i =  0}^{d-1}y_i\otimes\eta^i$,
and we set  \[ \first{\beta}{\eta}=y_0. \]
In particular, back to $A_\ell$, if we write
$\alpha_\ell = \sum_{i =  0}^{a-1}x_i\otimes\zeta_{\ell}^i$ where $a=\nu(\ell)$,
it is shown in~\cite{Allombert02} that $x_0=\first{\alpha_\ell}{\zeta_\ell}$ is a generating element for the extension
$\F_{p^\ell}/\F_{p}$.

\begin{proposition}
\label{depend}
With the notations above, there are precisely $\ell$ elements $x\in A_\ell$ that are solutions of~\eqref{h90} for $\zeta_\ell$
and satisfy $x^\ell=1\otimes\a_\ell$, namely, these are the $(1\otimes\zeta_\ell)^u\alpha_\ell=(\sigma^u\otimes 1)(\alpha_\ell)$
for $0\leq u<\ell$.
The corresponding generating elements for $\F_{p^\ell}/\F_{p}$ are the $\first{(\sigma^u\otimes 1)(\alpha_\ell)}{\zeta_\ell}=\sigma^u(x_0)$.
They all have the same minimal polynomial, which is a generating polynomial for $\F_{p^\ell}/\F_{p}$ that depends only on $\a_\ell$.
\end{proposition}
\begin{proof}
Clear.
\end{proof}

Now we consider Kummer algebras of various degrees.
Since we assumed the fields of scalars are defined from a system $\system$
of compatible roots of unity,
they are compatibly embedded:
for $\ell\,|\,m$ prime to $p$, we have the embedding
\[
\begin{array}{cccc}
  \embedcyc{\ell}{m}: & \F_{p}(\zeta_\ell) & \hookrightarrow & \F_{p}(\zeta_m) \\
  & \zeta_\ell & \mapsto & (\zeta_{m})^{\frac{m}{\ell}}.
\end{array}
\]

It is easily shown that, as an $\F_{p}$-algebra, $A_\ell$ is isomorphic to a product of copies of $\F_{p^\ell}(\zeta_\ell)$,
and $A_m$ to a product of copies of $\F_{p^m}(\zeta_m)$.
This allows to describe all $\F_{p}$-algebra morphisms from $A_\ell$ to $A_m$. However here we will focus only on a certain
subclass of them:
\begin{definition}
\label{Kembedding}
A \emph{Kummer embedding} of $A_\ell$ into $A_m$ is an \emph{injective} $\F_{p}$-algebra morphism $\Phi:A_\ell\hookrightarrow A_m$
such that:
\begin{itemize}
\item $\Phi$ extends the scalar embedding $1\otimes\embedcyc{\ell}{m}$
\item $\Phi$ commutes with $\sigma\otimes1$.
\end{itemize}
\end{definition}

\begin{proposition}
\label{Phialpha}
Let $\alpha_\ell\in A_\ell$ be a nonzero solution of~\eqref{h90} for $\zeta_\ell$, with Kummer constant $\a_\ell$.
Then, there is a $1$-to-$1$ correspondence between Kummer embeddings $\Phi:A_\ell\hookrightarrow A_m$ and solutions $\hat{\alpha}\in A_m$
of~\eqref{h90} for $(\zeta_m)^{\frac{m}{\ell}}$ that satisfy $(\hat{\alpha})^\ell=1\otimes\embedcyc{\ell}{m}(\a_\ell)$,
given by \[ \Phi\quad\longleftrightarrow\quad\hat{\alpha}=\Phi(\alpha_\ell). \]
\end{proposition}
\begin{proof}
Direct consequence of Proposition~\ref{alphagen} and Definition~\ref{Kembedding}.
\end{proof}

Actually, Kummer embeddings are easily characterized:
\begin{proposition}
\label{Phipsi}
There is a natural $1$-to-$1$ correspondence between Kummer embeddings $\Phi:A_\ell\hookrightarrow A_m$
and embeddings of finite fields $\psi:\F_{p^\ell}\hookrightarrow\F_{p^m}$, given by:
\begin{itemize}
\item If $\Phi$ is a Kummer embedding, then $\Phi$ maps $\F_{p^\ell}\otimes1$ into $\F_{p^m}\otimes 1$.
Thus the restriction of $\Phi$ to $\F_{p^\ell}\otimes1$ is of the form $\psi\otimes1$ for some $\psi:\F_{p^\ell}\hookrightarrow\F_{p^m}$.
\item Conversely, if $\psi:\F_{p^\ell}\hookrightarrow\F_{p^m}$ is an embedding of finite fields, then $\Phi=\psi\otimes\embedcyc{\ell}{m}$
is a Kummer embedding.
\end{itemize}
Moreover, this correspondence commutes with composition of embeddings.
\end{proposition}
\begin{proof}
Let $\Phi$ be a Kummer embedding. Being a $\F_{p}$-algebra morphism, it satisfies $\Phi(\beta^p)=\Phi(\beta)^p$ for all $\beta\in A_\ell$.
By Lemma~\ref{FrobFrob}, this means that $\Phi$ commutes with $\sigma\otimes\sigma$, and thus,
also with $(\sigma\otimes 1)^{-1}\circ(\sigma\otimes\sigma)=1\otimes\sigma$.
This gives that $\Phi$ maps $\F_{p^\ell}\otimes1$ into $\F_{p^m}\otimes 1$.
The other assertions are clear.
\end{proof}

\begin{corollary}
\label{alphapsi}
Let $\alpha_\ell\in A_\ell$ be a nonzero solution of~\eqref{h90} for $\zeta_\ell$,
with Kummer constant $\a_\ell$, and
let $\hat{\alpha}\in A_m$ be a solution of~\eqref{h90} for $(\zeta_m)^{\frac{m}{\ell}}$
that satisfies $(\hat{\alpha})^\ell=1\otimes\embedcyc{\ell}{m}(\a_\ell)$.
Then:
\begin{itemize}
\item $\hat{\alpha}\in\F_{p^m}\otimes\F_{p}((\zeta_m)^{\frac{m}{\ell}})\;\subset\;A_m$
\item the assignation $\first{\alpha_{\ell}}{\zeta_\ell}\mapsto\first{\hat{\alpha}}{(\zeta_m)^{\frac{m}{\ell}}}$
defines an embedding $\psi:\F_{p^\ell}\hookrightarrow\F_{p^m}$
\item $\Phi=\psi\otimes\embedcyc{\ell}{m}$ is the unique Kummer embedding such that $\Phi(\alpha_\ell)=\hat{\alpha}$.
\end{itemize}
\end{corollary}
\begin{proof}
By Proposition~\ref{Phialpha} there is a unique Kummer embedding $\Phi$ such that $\Phi(\alpha_\ell)=\hat{\alpha}$.
By Proposition~\ref{Phipsi} we have $\Phi=\psi\otimes\embedcyc{\ell}{m}$ for some $\psi:\F_{p^\ell}\hookrightarrow\F_{p^m}$.
Writing $\alpha_\ell=\sum_{i=0}^{a-1}x_i\otimes\zeta_{\ell}^i$,
it follows \[ \hat{\alpha}=\Phi(\alpha_\ell)=\sum_{i=0}^{a-1}\psi(x_i)\otimes(\zeta_m)^{\frac{mi}{\ell}}. \]
Thus $\first{\hat{\alpha}}{(\zeta_m)^{\frac{m}{\ell}}}=\psi(x_0)=\psi(\first{\alpha_{\ell}}{\zeta_\ell})$,
and since $\first{\alpha_{\ell}}{\zeta_\ell}$ generates $\F_{p^\ell}$, this uniquely characterizes $\psi$.
\end{proof}

We can now state and prove Allombert's algorithm.\footnote{Actually,
  Allombert's algorithm does not require $A_m$,
  but only the smaller algebra $\F_{p^m}\otimes\F_p(\zeta_l)$.
  This will make no difference for us.}

\begin{algorithm}
  \caption{(Allombert's)}
  \label{algo:allombert}
  \begin{algorithmic}[1]
    \REQUIRE {blackbox $\F_{p^\ell}, \F_{p^m}$, for $\ell\,|\,m$ integers prime to $p$.}
    \ENSURE {$s\in\F_{p^\ell}, t\in\F_{p^m}$, such that the assignation $s\mapsto t$
      defines an embedding $\psi:\F_{p^\ell}\hookrightarrow\F_{p^m}$.}
  \STATE Prepare the Kummer algebras $A_\ell$ and $A_m$.
  \STATE Find $\alpha_\ell\in A_\ell$, $\alpha_m\in A_m$ nonzero solutions of \eqref{h90} for $\zeta_\ell$
  and $\zeta_m$ respectively.
  \STATE Compute their Kummer constants: $(\alpha_\ell)^\ell=1\otimes\a_\ell$, 
  $(\alpha_m)^m=1\otimes\a_m$.
  \STATE Compute $\kappa$ a $\ell$-th root of $\embedcyc{\ell}{m}(\a_\ell)/\a_m$.
  \STATE Return $\first{\alpha_{\ell}}{\zeta_\ell}$ and $\first{(1\otimes\kappa)(\alpha_m)^{\frac{m}{\ell}}}{(\zeta_m)^{\frac{m}{\ell}}}$.
  \end{algorithmic}
\end{algorithm}
\begin{proposition}
  Algorithm~\ref{algo:allombert} is correct: it returns elements that define an
  embedding $\psi:\F_{p^\ell}\hookrightarrow\F_{p^m}$.
\end{proposition}
\begin{proof}
By Propositions~\ref{Phialpha} and~\ref{Phipsi}, there exists $\hat{\alpha}\in A_m$
solution of~\eqref{h90} for $(\zeta_m)^{\frac{m}{\ell}}$ that satisfy $(\hat{\alpha})^\ell=1\otimes\embedcyc{\ell}{m}(\a_\ell)$.
On the other hand, $(\alpha_m)^{\frac{m}{\ell}}\in A_m$ is also a solution of~\eqref{h90} for $(\zeta_m)^{\frac{m}{\ell}}$,
thus $\hat{\alpha}=(1\otimes\lambda)(\alpha_m)^{\frac{m}{\ell}}$ for some $\lambda\in\F_{p}(\zeta_m)$.
It follows $\embedcyc{\ell}{m}(\a_\ell)/\a_m=\lambda^\ell$ is a $\ell$-th power,
and $\kappa = (\zeta_{m})^{\frac{um}{\ell}}\lambda$ for some integer $u$.
Now we can replace $\hat{\alpha}$
with $(1\otimes(\zeta_{m})^{\frac{um}{\ell}})(\hat{\alpha})=(1\otimes\kappa)(\alpha_m)^{\frac{m}{\ell}}$
and conclude with Corollary~\ref{alphapsi}.
\end{proof}
From this proof and Proposition~\ref{depend}, we remark that another choice of the $l$-th root $\kappa$
only changes $\psi$ by a power of $\sigma$.
    
\section{Standard solutions of (H90)}
\label{sec:compatibleH90}

There are two points in Algorithm~\ref{algo:allombert} on which we would like to improve:
\begin{description}
\item[\emph{Uniqueness:}] As mentioned, the element $\first{\alpha_{\ell}}{\zeta_\ell}$ is a generating element for $\F_{p^{\ell}}$,
or equivalently, it provides a defining irreducible polynomial of degree $\ell$.
However this polynomial depends on the choice of $\alpha_{\ell}$
(even though only through~$\a_\ell$, cf. Proposition~\ref{depend}).
\item[\emph{Compatibility:}] The embedding $\psi$ depends on a constant $\kappa$,
which itself depends on the choice of $\alpha_\ell,\alpha_m$ (and of a $\ell$-th root extraction).
Thus, given a certain number of finite fields, in order to ensure compatibility of the various
embeddings between them, one has to keep track of these constants $\kappa$
for all pairs $(\ell,m)$,
which grows \emph{quadratically} with the number of fields.
\end{description}

It would be nice if one could force $\kappa=1$, that is, if $\alpha_\ell,\alpha_m$
and $\Phi:A_\ell\hookrightarrow A_m$
could be chosen so that $\Phi(\alpha_\ell)=(\alpha_m)^{\frac{m}{\ell}}$.
From the description of Algorithm~\ref{algo:allombert},
this requires $\a_m=\embedcyc{\ell}{m}(\a_\ell)$.
Thus,
necessarily $\a_m$ lies in the subfield $\F_{p}((\zeta_m)^{\frac{m}{\ell}})$ of $\F_{p}(\zeta_m)$.
Possibly this condition could fail if $A_\ell$ and $A_m$ do not have the same field of scalars.
This motivates:
\begin{definition}
\label{complete}
A Kummer algebra is \emph{complete} if it is of the largest degree for a given level.
\end{definition}
Thus, the complete Kummer algebra of level $a$ is
\[ A_{p^a-1}=\F_{p^{p^a-1}}\otimes\F_{p^a} \]
with field of scalars $\F_{p^a}=\F_{p}(\zeta_{p^a-1})$ given by the corresponding $\zeta_{p^a-1}$ in our compatible system $\system$,
e.g. defined by a (pseudo-)Conway polynomial of degree $a$.
\begin{lemma}
\label{Kummer_bizarre}
All nonzero solutions $\alpha_{p^a-1}\in A_{p^a-1}$ of \eqref{h90} for $\zeta_{p^a-1}$,
have the same Kummer constant $\a_{p^a-1}=(\zeta_{p^a-1})^a$.
\end{lemma}
\begin{proof}
From Lemma~\ref{FrobFrob} and the fact that $\sigma^a$ is trivial on $\F_{p}(\zeta_{p^a-1})\simeq\F_{p^a}$ we get
\begin{equation*}
\begin{split}
(\alpha_{p^a-1})^{p^a}=(\sigma^a\otimes\sigma^a)(\alpha_{p^a-1})&=(\sigma^a\otimes1)(\alpha_{p^a-1})\\
&=(1\otimes\zeta_{p^a-1})^a\alpha_{p^a-1}.
\end{split}
\end{equation*}
We conclude since $\alpha_{p^a-1}$ is invertible.
\end{proof}
\begin{definition}
\label{alphastandard}
Let $l$ be an integer prime by $p$.
We define the \emph{standard} Kummer constant of order $\ell$ as
\[ \stda_{\ell}=(\embedcyc{l}{p^a-1})^{-1}((\zeta_{p^a-1})^a)\in\F_p(\zeta_\ell) \]
where $a=\nu(\ell)$ is the level of $A_\ell$.

We say a solution $\alpha_\ell\in A_\ell$ of~\eqref{h90} for $\zeta_\ell$ is
standard if its Kummer constant is standard:
\[ (\alpha_\ell)^\ell=1\otimes\stda_{\ell}. \]
Then, by a \emph{decorated} Kummer algebra we mean a pair
\[ (A_\ell,\alpha_{\ell}) \]
with such $\alpha_\ell$ standard.
\end{definition}
Observe that $\embedcyc{l}{p^a-1}$ is an isomorphism when $a=\nu(\ell)$,
so $\stda_\ell$ is well defined.

%%%%%%%%   Supprimable si redondant avec la section 6 ou 7  %%%%%%%%%%
It seems that, in order to compute $\stda_\ell$
so as to check that $\alpha_\ell$ is standard, one has to
consider the primitive root $\zeta_{p^a-1}$ of order $p^a-1$ in our
system $\system$.
When $\system$ is constructed from (pseudo-)Conway polynomials,
one already needs $\zeta_{p^a-1}$ in order to construct $\zeta_\ell$.
However for a more general system $\system$, one would like to avoid this cost
as much as possible.
Actually, what we really need is only
precompute $\stda_\ell$ in $\F_p(\zeta_\ell)$,
and maybe one could devise more efficient ways to deal with this.
To start with, observe it can also be written
$\stda_\ell=(\embedcyc{l}{\frac{p^a-1}{d}})^{-1}((\zeta_{\frac{p^a-1}{d}})^{\frac{a}{d}})$,
where $d=\gcd(a,p^a-1)$, so we only need go to order $\frac{p^a-1}{d}$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For complete algebras, Lemma~\ref{Kummer_bizarre} asserts that all nonzero $\alpha_{p^a-1}$ are standard.
\begin{proposition}
\label{standardexiste}
Let $l$ be an integer not divisible by $p$.
Then $A_\ell$ can be decorated, i.e. it admits a standard $\alpha_\ell$.
Moreover, this $\alpha_\ell$ is unique up to a $\ell$-th root of unity.
\end{proposition}
\begin{proof}
Let $\alpha'_\ell$ be any nonzero solution of~\eqref{h90} for $\zeta_\ell$.
Set $a=\nu(\ell)$, pick any $\alpha_{p^a-1}\in A_{p^a-1}$ standard (Lemma~\ref{Kummer_bizarre}),
and pick any Kummer embedding $\Phi:A_\ell\hookrightarrow A_{p^a-1}$ (Proposition~\ref{Phipsi}).
Then $\Phi(\alpha'_\ell)$ and $(\alpha_{p^a-1})^{\frac{p^a-1}{\ell}}$ are two nonzero solutions of~\eqref{h90}
for $\embedcyc{l}{p^a-1}(\zeta_\ell)=(\zeta_{p^a-1})^{\frac{p^a-1}{\ell}}$ in $A_{p^a-1}$, thus there is a scalar $\lambda\in\F_{p}(\zeta_{p^a-1})$
such that
\begin{equation*}
(\alpha_{p^a-1})^{\frac{p^a-1}{\ell}}=(1\otimes\lambda)\Phi(\alpha'_\ell)=\Phi((1\otimes\tilde\lambda)\alpha'_\ell).
\end{equation*}
where $\tilde\lambda=(\embedcyc{l}{p^a-1})^{-1}(\lambda)\in\F_{p}(\zeta_{\ell})$.

Setting $\alpha_\ell=(1\otimes\eta\tilde\lambda)\alpha'_\ell\in A_\ell$ for $\eta\in\F_{p}(\zeta_\ell)$,
we get $\a_{\ell}=\eta^\ell(\embedcyc{l}{p^a-1})^{-1}(\stda_{p^a-1})=\eta^\ell\stda_\ell$, and
thus the standard $\alpha_\ell\in A_\ell$ are the $(1\otimes\zeta_\ell^u\tilde\lambda)\alpha'_\ell$, for $0\leq u<\ell$.
\end{proof}
This allows:

\begin{algorithm}
  \caption{(decoration-standardization)}
  \label{algo:decoration}
  \begin{algorithmic}[1]
    \REQUIRE {blackbox $\F_{p^\ell}$, for $\ell$ an integer prime to $p$.}
    \ENSURE {$(A_\ell,\alpha_\ell)$ decorated, $P_\ell$ irreducible polynomial of degree $\ell$,
      and $s\in\F_{p^\ell}$ a generating element inducing $\F_{p^\ell}\simeq\F_{p}[T]/(P_\ell)$.}
  \STATE Prepare the Kummer algebra $A_\ell$.
  \STATE Prepare $\stda_\ell=(\embedcyc{l}{p^a-1})^{-1}((\zeta_{p^a-1})^a)\in\F_p(\zeta_\ell)$.
  \STATE Find $\alpha'_\ell\in A_\ell$ nonzero solution of \eqref{h90} for $\zeta_\ell$.
  \STATE Compute its Kummer constant: $(\alpha'_\ell)^\ell=1\otimes\a'_\ell$.
  \STATE Compute $\kappa$ a $\ell$-th root of $\stda_\ell/\a'_\ell$.
  \STATE Set $\alpha_{\ell}=(1\otimes\kappa)\alpha'_\ell$.
  \STATE Compute $P_\ell$ the minimal polynomial of $\first{\alpha_{\ell}}{\zeta_\ell}$ over $\F_{p}$.
  \STATE Return $(A_\ell,\alpha_\ell)$, $P_\ell$, and $\first{\alpha_\ell}{\zeta_\ell}$.
  \end{algorithmic}
\end{algorithm}

Algorithm~\ref{algo:decoration} is correct: indeed Proposition~\ref{standardexiste} ensures that
a standard $\alpha_{\ell}$ exists, and thus, $\alpha'_\ell=(1\otimes\kappa^{-1})\alpha_{\ell}$
for some $\kappa\in\F_{p}(\zeta_\ell)$, so $\stda_\ell/\a'_\ell=\kappa^\ell$ is a $\ell$-th power.

By Proposition~\ref{depend}, we also note that $P_\ell$ only depends on $\stda_{\ell}$,
and thus it only depends on the given system $\system$ of compatible roots,
possibly up to order $p^{\nu(\ell)}-1$
(e.g. $P_\ell$ only depends the Conway polynomial of degree $\nu(\ell)$).
In this sense, $P_\ell$ gives a standard definition of $\F_{p^\ell}$.

By design, decorated Kummer algebras of the same level admit standard Kummer embeddings,
under which the corresponding standard solutions of~\eqref{h90} are power-compatible:
\begin{proposition}
\label{embedincomplete}
Let $\ell\,|\,m$ be integers prime to $p$ and such $\nu(\ell)=\nu(m)=a$.
Let $(A_\ell,\alpha_\ell)$, $(A_m,\alpha_m)$ be decorated Kummer algebras
of degree $\ell,m$ respectively, and of the same level $a$.
Then, there is a unique Kummer embedding
\[ \embedalg{\ell}{m}:A_\ell\hookrightarrow A_m \]
such that $\embedalg{\ell}{m}(\alpha_\ell)=(\alpha_m)^{\frac{m}{\ell}}$.
\end{proposition}
\begin{proof}
Proposition~\ref{Phialpha} with $\hat{\alpha}=(\alpha_m)^{\frac{m}{\ell}}$.
\end{proof}
Most often we will apply Proposition~\ref{embedincomplete} with $m=p^a-1$.

On the other hand, 
since power-compatibility implies $\a_m=\embedcyc{\ell}{m}(\a_\ell)$,
it cannot be satisfied for a Kummer embedding
between decorated Kummer algebras of different levels.
However, at least between \emph{complete} decorated algebras, we can request some \emph{norm}-compatibility
instead.

Let $A_m$ be a Kummer algebra of level $b=\nu(m)$, so
\[ A_m=\F_{p^m}\otimes\F_{p}(\zeta_m)\overset{\sim}{\longrightarrow}\F_{p^m}\otimes\F_{p}(\zeta_{p^b-1}) \]
where the isomorphism is given by $1\otimes\embedcyc{m}{p^b-1}$.
Then, for an integer $a\,|\,b$,
the subalgebra of $A_m$ invariant under $1\otimes\sigma^a$ is identified
by this isomorphism with
\[ (A_m)^{1\otimes\sigma^a}\simeq\F_{p^m}\otimes\F_{p}((\zeta_{p^b-1})^{\frac{p^b-1}{p^a-1}}) \]
where $(\zeta_{p^b-1})^{\frac{p^b-1}{p^a-1}}=\operatorname{N}_{\F_{p^b}/\F_{p^a}}(\zeta_{p^b-1})=\embedcyc{p^a-1}{p^b-1}(\zeta_{p^a-1})$.

\begin{definition}
\label{def:norm}
Given a Kummer algebra $A_n$, and integers $a\,|\,b\,|\,\nu(n)$, we define
the \emph{scalar norm} operator
\[
\begin{array}{cccc}
  \norm_{b/a,A_n}: & (A_n)^{1\otimes\sigma^b} & \to & (A_n)^{1\otimes\sigma^a} \\
  & \gamma & \mapsto & \prod_{0\leq j<\frac{b}{a}} (1 \otimes \sigma^{ja})(\gamma).
\end{array}
\]
\end{definition}
This is well-defined, i.e. the image of $\norm_{b/a,A_n}$ is invariant under $1\otimes\sigma^a$ as specified.
Often the ambiant algebra $A_n$ will be implicit,
and we will write $\norm_{b/a}$ instead of $\norm_{b/a,A_n}$.

By construction, $\norm_{b/a}$ acts on $1\otimes\F_{p^b}^\times$ as $1\otimes\operatorname{N}_{\F_{p^b}/\F_{p^a}}$.
They are multiplicative:
\[
  \norm_{b/a}(\gamma\gamma')=\norm_{b/a}(\gamma)\norm_{b/a}(\gamma'),
\]
transitive: \[ \norm_{c/a} = \norm_{b/a}\circ\norm_{c/b},\] and they
commute with $\sigma\otimes1$.
\begin{proposition}
\label{embedcomplete}
Let $a\,|\,b$ be integers,
and $(A_{p^a-1},\alpha_{p^a-1})$, $(A_{p^b-1},\alpha_{p^b-1})$ decorated complete Kummer algebras
of level $a,b$ respectively.
Then there is a unique Kummer embedding
\[ \embedalg{p^a-1}{p^b-1}:A_{p^a-1}\hookrightarrow A_{p^b-1} \]
such that $\embedalg{p^a-1}{p^b-1}(\alpha_{p^a-1})=\norm_{b/a}(\alpha_{p^b-1})$.
\end{proposition}
\begin{proof}
By Lemma~\ref{FrobFrob} and the properties of the norm,
\begin{equation*}
\begin{split}
(\norm_{b/a}(\alpha_{p^b-1}))^{p^a}&=(\sigma^a\otimes\sigma^a)(\norm_{b/a}(\alpha_{p^b-1}))\\
&=(\sigma^a\otimes1)(\norm_{b/a}(\alpha_{p^b-1}))\\
&=\norm_{b/a}((\sigma^a\otimes1)(\alpha_{p^b-1}))\\
&=\norm_{b/a}((1\otimes(\zeta_{p^b-1})^a)\alpha_{p^b-1})\\
&=(1\otimes\embedcyc{p^a-1}{p^b-1}(\zeta_{p^a-1})^a)\norm_{b/a}(\alpha_{p^b-1}).
\end{split}
\end{equation*}
So $\hat\alpha=\norm_{b/a}(\alpha_{p^b-1})$
satisfies $(\hat{\alpha})^{p^a-1}=1\otimes\embedcyc{p^a-1}{p^b-1}(\a_{p^a-1})$
and we conclude with Proposition~\ref{Phialpha}.
\end{proof}



\section{Standard embeddings}
\label{sec:construction}

In Proposition~\ref{embedincomplete}, we saw how to construct a standard
power-compatible embedding of a decorated Kummer algebra into
its decorated complete algebra,
and in Proposition~\ref{embedcomplete}, a standard
norm-compatible embedding between decorated complete algebras
of dividing levels.
 
Now consider general $\ell\,|\,m$ not divisible by $p$,
set $a=\nu(\ell)$, $b=\nu(m)$, and consider the diagram
\begin{equation*}
\label{3cotes}
\begin{CD}
(A_{p^a-1},\alpha_{p^a-1}) @>{\embedalg{p^a-1}{p^b-1}}>> (A_{p^b-1},\alpha_{p^b-1}) \\
@A{\embedalg{\ell}{p^a-1}}AA @AA{\embedalg{m}{p^b-1}}A\\
(A_\ell,\alpha_\ell) @. (A_m,\alpha_m)
\end{CD}
\end{equation*}
of standard embeddings of decorated algebras.
\begin{lemma}
In this setting, there exists a unique Kummer embedding
\[ \embedalg{\ell}{m}:A_\ell\hookrightarrow A_m \]
that makes the diagram commute.
\end{lemma}
\begin{proof}
Consider $\hat{\hat\alpha}=\embedalg{p^a-1}{p^b-1}(\embedalg{\ell}{p^a-1}(\alpha_\ell))\in A_{p^b-1}$.
Then $\hat{\hat\alpha}$ is invariant under $\sigma^\ell\otimes1$ and
under $1\otimes\sigma^a$ (because $\alpha_\ell$ is),
thus, a fortiori, invariant under $\sigma^m\otimes1$ and
under $1\otimes\sigma^b$, which means it lies in the image
of $\embedalg{m}{p^b-1}$.
We can then set $\hat\alpha=(\embedalg{m}{p^b-1})^{-1}(\hat{\hat\alpha})$.
If $\embedalg{\ell}{m}$ exists, then necessarily it maps $\alpha_\ell$
to $\hat\alpha$.
However, chasing in the diagram, it is easily seen that $\hat\alpha$ is
a solution of~\eqref{h90} for $(\zeta_m)^{\frac{m}{\ell}}$ that
satisfies $(\hat{\alpha})^\ell=1\otimes\embedcyc{\ell}{m}(\a_\ell)$,
and we conclude with Proposition~\ref{Phialpha}.
\end{proof}
This existence result is ``constructive'', but impractical,
since it requires computations in the possibly very large algebra $A_{p^b-1}$.
However, as in Algorithm~\ref{algo:allombert}, one should be able
to write $\hat{\alpha}=(1\otimes\kappa)(\alpha_m)^{\frac{m}{\ell}}$
for some $\kappa\in\F_{p}(\zeta_m)$. Moreover $\hat{\alpha}$ is uniquely
determined by our data, thus, so should $\kappa$.
Now our aim is to give an explicit expression for this $\kappa=\kappa_{\ell,m}$.
We start with the case of complete algebras.
\begin{proposition}
\label{key}
In the complete algebra $A_{p^b-1}$ we have
\[ (\alpha_{p^b-1})^{\frac{p^b-1}{p^a-1}}=(1\otimes\zeta_{p^b-1})^{\frac{(b-a)p^{b+a}-bp^b+ap^a}{(p^a-1)^2}}\norm_{b/a}(\alpha_{p^b-1}).\]
\end{proposition}
\begin{proof}
Using first Lemma~\ref{FrobFrob}, and then~\eqref{h90}, we get:
\begin{equation*}
\begin{split}
\frac{(\alpha_{p^b-1})^{\frac{p^b-1}{p^a-1}}}{\norm_{b/a}(\alpha_{p^b-1})}&=\prod_{0\leq j<\frac{b}{a}}\frac{(\sigma^{ja}\otimes\sigma^{ja})(\alpha_{p^b-1})}{(1\otimes\sigma^{ja})(\alpha_{p^b-1})}\\
&=\prod_{0\leq j<\frac{b}{a}}(1\otimes\sigma^{ja})\left(\frac{(\sigma^{ja}\otimes 1)(\alpha_{p^b-1})}{\alpha_{p^b-1}}\right)\\
&=\prod_{0\leq j<\frac{b}{a}}(1\otimes\sigma^{ja})(1\otimes\zeta_{p^b-1})^{ja}\\
&=(1\otimes\zeta_{p^b-1})^{\sum_{0\leq j<\frac{b}{a}}jap^{ja}}.
\end{split}
\end{equation*}
We conclude thanks to the identity
\[ \sum_{0\leq j<n}jT^j=T\frac{d}{dT}\!\left(\frac{T^n-1}{T-1}\right)=\frac{(n-1)T^{n+1}-nT^n+T}{(T-1)^2}. \]
\end{proof}
\begin{corollary}
\label{explicit_general_standard_embedding}
Let $(A_\ell,\alpha_\ell)$ and $(A_m,\alpha_m)$ be decorated Kummer algebras,
of respective degrees $\ell\,|\,m$ prime to $p$.
Then the standard Kummer embedding $\embedalg{\ell}{m}:A_\ell\hookrightarrow A_m$
is defined by the assignation
$\alpha_\ell\;\mapsto\;(1\otimes\kappa_{\ell,m})(\alpha_m)^{\frac{m}{\ell}}$, where
$\kappa_{\ell,m}=(\embedcyc{m}{p^b-1})^{-1}((\zeta_{p^b-1})^{-\frac{(b-a)p^{b+a}-bp^b+ap^a}{(p^a-1)\ell}})$.
\end{corollary}
\begin{proof}
It suffices to check that $\embedalg{p^a-1}{p^b-1}(\embedalg{\ell}{p^a-1}(\alpha_\ell))$
and the image of the right-hand-side under $\embedalg{m}{p^b-1}$ coincide in $A_{p^b-1}$.
However, we have $\embedalg{p^a-1}{p^b-1}(\embedalg{\ell}{p^a-1}(\alpha_\ell))=\norm_{b/a}(\alpha_{p^b-1})^{\frac{p^a-1}{\ell}}$,
while $\embedalg{m}{p^b-1}((\alpha_m)^{\frac{m}{\ell}})=(\alpha_{p^b-1})^{\frac{p^b-1}{\ell}}$,
and we conclude with Proposition~\ref{key}
\end{proof}
%%%%%%%    Supprimable si redondant avec la section 6 ou 7 %%%%%%%%%%%%%%
As in the discussion following Defintion~\ref{alphastandard}, observe
that the computation of $\kappa_{\ell,m}$ possibly requires a root of order
up to $p^b-1$ in $\system$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{proposition}
\label{standard_K_embeddings_compatibles}
Standard Kummer embeddings are compatible with composition:
if $(A_\ell,\alpha_\ell)$, $(A_m,\alpha_m)$, and $(A_n,\alpha_n)$ are decorated Kummer algebras
with $\ell\,|\,m\,|\,n$, the corresponding standard embeddings
satisfy $\embedalg{\ell}{n}=\embedalg{m}{n}\circ\embedalg{\ell}{m}$.
\end{proposition}
\begin{proof}
We have to show $\embedalg{\ell}{n}(\alpha_\ell)=\embedalg{m}{n}(\embedalg{\ell}{m}(\alpha_\ell))$:
nothing but a pleasant computation with the explicit constants given 
by Corollary~\ref{explicit_general_standard_embedding}.

Alternatively, set $a=\nu(\ell),b=\nu(m),c=\nu(n)$, and decorate $A_{p^a-1},A_{p^b-1},A_{p^c-1}$.
It suffices to show that $\embedalg{\ell}{n}(\alpha_\ell)$
and $\embedalg{m}{n}(\embedalg{\ell}{m}(\alpha_\ell))$ have the same image under $\embedalg{n}{p^c-1}$
in $A_{p^c-1}$.
Chasing in the diagram
\begin{equation*}
\begin{CD}
A_{p^a-1} @>>> A_{p^b-1} @>>> A_{p^c-1} \\
@AAA @AAA @AAA\\
A_\ell @>>> A_m @>>> A_n
\end{CD}
\end{equation*}
we see indeed this common image
is $\norm_{c/a}(\alpha_{p^c-1})^{\frac{p^a-1}{\ell}}$.
\end{proof}

By a \emph{decorated finite field} (of degree $\ell$, an integer prime to $p$),
we mean a pair $(\F_{p^\ell},\alpha_\ell)$,
where $\F_{p^\ell}$ is a (blackbox) finite field,
and $\alpha_\ell$ a standard solution of~\eqref{h90} in the corresponding
Kummer algebra $A_\ell=\F_{p^\ell}\otimes\F_{p}(\zeta_\ell)$.

If $(\F_{p^\ell},\alpha_\ell)$ is a decorated finite field,
then from $\alpha_\ell$ one derives the standard generating
element $x_0=\first{\alpha_\ell}{\zeta_\ell}$ of $\F_{p^\ell}$.
Conversely, given such a standard $x_0$, one can retrieve $\alpha_\ell$,
e.g. using the equations given in~\cite{Allombert02}.

We can finally state:
\begin{algorithm}
  \caption{(standard, compatible Allombert)}
  \label{algo:std_embed}
  \begin{algorithmic}[1]
    \REQUIRE {decorated finite fields $(\F_{p^\ell},\alpha_\ell)$, $(\F_{p^m},\alpha_m)$, for $\ell\,|\,m$ integers prime to $p$.}
    \ENSURE {$s\in\F_{p^\ell}, t\in\F_{p^m}$, such that the assignation $s\mapsto t$
      defines a standard embedding $\embed{\ell}{m}:\F_{p^\ell}\hookrightarrow\F_{p^m}$,
      compatible with composition.}
  \STATE Compute $\kappa_{\ell,m}=(\embedcyc{m}{p^b-1})^{-1}((\zeta_{p^b-1})^{-\frac{(b-a)p^{b+a}-bp^b+ap^a}{(p^a-1)\ell}})$ where $a=\nu(\ell)$, $b=\nu(m)$.
  \STATE Return $\first{\alpha_\ell}{\zeta_\ell}$ and $\first{(1\otimes\kappa)(\alpha_m)^{\frac{m}{\ell}}}{(\zeta_m)^{\frac{m}{\ell}}}$.
  \end{algorithmic}
\end{algorithm}
\begin{proposition}
\label{standard_ff_embeddings_compatibles}
Standard finite field embeddings are compatible with composition:
if $(\F_{p^\ell},\alpha_\ell)$, $(\F_{p^m},\alpha_m)$, and $(\F_{p^n},\alpha_n)$ are decorated finite fields
with $\ell\,|\,m\,|\,n$, the corresponding standard embeddings
satisfy $\embed{\ell}{n}=\embed{m}{n}\circ\embed{\ell}{m}$.
\end{proposition}
\begin{proof}
Corollary~\ref{Phipsi} and Proposition~\ref{standard_K_embeddings_compatibles}
\end{proof}

\section{Implementation}
\label{sec:implementation}

To demonstrate that our strategy is practical, we implemented it using the
Julia library Nemo~\cite{Fieker:2017:NCA:3087604.3087611} and the C library
Flint~\cite{flint}. Critical routines are written in
C\footnote{\url{https://github.com/erou/gf-h90-lattice/tree/master/implementation}}
for performance, whereas the high level tasks are performed in Julia. Our code is available as a Julia
package\footnote{\url{https://github.com/erou/LatticeGFH90.jl}}. Our
implementation is limited to word-sized characteristics. Since Algorithm~\ref{algo:std_embed} require decorated finite
fields, we first measure the time needed by Algorithm~\ref{algo:decoration} to
perform this task. In fact, the computations in Algorithm~\ref{algo:std_embed}
take negligible time in comparison with the ones in
Algorithm~\ref{algo:decoration}. The bottleneck of the decoration is the root
extraction: this takes almost all the time.

\section{Conclusion and future work}
\label{sec:conclusion}

\paragraph{Acknowledgments}
We would like to thank Éric Schost for fruitful discussions and for
helping bootstrap this work during a visit by two of the authors to
the University of Waterloo. %
We acknowledge financial support from the French ANR-15-CE39-0013
project \emph{Manta}, the \emph{OpenDreamKit} Horizon 2020 European
Research Infrastructures project (\#676541), and from the French
Domaine d'Intérêt Majeur \emph{Math'Innov}.


\bibliographystyle{plain}
\bibliography{gf-h90}

\end{document}

% LocalWords:  embeddings computable Frobenius Lenstra cyclotomic
